Bayesian RKHS-based methods in functional regression

### 2022-06-08 slides/seio.tex
Changes in slides bis

\begin{frame}{Table of contents}
  \tableofcontents
\end{frame}

@@

\vspace{1em}
  Instead of assuming \(\beta \in L^2[0, 1]\), we consider \maroon{\(\beta \in \mathcal H(K)\)}, the RKHS associated with the covariance function \(K(t, s)=\mathbb E[X(t)X(s)]\) of the centered process \(X\). Moreover, we substitute \(\dotprod{X}{\beta}_2\) for \maroon{\(\dotprod{X}{\beta}_K\)}.
\end{frame}

@@

\begin{frame}{Parameter estimation overview}

  We have a parameter vector \(\theta = (p, \beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2)\), on which we impose a prior distribution \(\pi(\theta)\). In practice, we are establishing a prior distribution on the functional space \(\Hcal(K)\) via a density argument.

  Afterwards we apply \textit{Bayes'} theorem in the usual way to obtain the posterior distribution of the parameters given the data.

  \begin{block}{Bayes's theorem}
      \[
      \pi(\theta \mid \mathcal D_n) \propto \left( \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right)\pi(\theta).
      \]
  \end{block}
  \vspace{1em}

  Finally, we obtain samples of the approximate posterior distribution using an algorithm of the family of \textit{Markov chain Monte Carlo} (MCMC) methods.

@@

\begin{alertblock}{Hyperparameter selection}
  \vspace{0.1em}
  For \(b_0\) we use a computational approximation of the MLE; the rest of hyperparameters are set via \textit{cross-validation}.
\end{alertblock}
\metroset{block=fill}
\end{frame}

@@

\vspace{1em}
  \begin{block}{Logistic log-posterior}
      \vspace{-1em}
    \begin{align*}
  \log \pi(\theta \mid \mathcal D_n) \propto {} & \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi_{X_i}(\alpha)\right)Y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{X_i}(\alpha)\right\}\right)\right]\\
  \quad &+ \frac{1}{2}\log |G_\tau| - p\log \sigma -\frac{1}{2g\sigma^2} (b - b_0)'G_\tau(b - b_0).
\end{align*}
Remember that \(\Psi_{X_i}(\alpha) = \sum_j \beta_j X_i(t_j)\).
  \end{block}
\end{frame}

@@

  \begin{exampleblock}{The road ahead}
    \begin{itemize}
      \item Reduce the variance in predictions.
      \item Experiment with other prior distributions.
      \item Package the Python algorithm and make it compatible with standard libraries (e.g. \textit{sklearn}, \textit{scikit-fda}).
    \end{itemize}
  \end{exampleblock}

>>>>

% \begin{frame}{Table of contents}
%   \tableofcontents
% \end{frame}

@@

\vspace{1em}
  Instead of assuming \(\beta \in L^2[0, 1]\), we consider \maroon{\(\beta \in \mathcal H(K)\)}, the RKHS associated with the covariance function \(K(t, s)=\mathbb E[X(t)X(s)]\) of the centered process \(X\) (i.e. \(\E [X(t)]=0\) for all \(t\)). Moreover, we substitute \(\dotprod{X}{\beta}_2\) for \maroon{\(\dotprod{X}{\beta}_K\)}.
\end{frame}

@@

% \begin{frame}{Parameter estimation overview}
%
%   We have a parameter vector \(\theta = (p, \beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2)\), on which we impose a prior distribution \(\pi(\theta)\). In practice, we are establishing a prior distribution on the functional space \(\Hcal(K)\) via a density argument.
%
%   \vspace{1em}
%
%   Afterwards we apply \textit{Bayes'} theorem in the usual way to obtain the posterior distribution of the parameters given the data.
%
%   \begin{block}{Bayes's theorem}
%       \[
%       \pi(\theta \mid \mathcal D_n) \propto \left( \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right)\pi(\theta).
%       \]
%   \end{block}
%   \vspace{1em}
%
%   Finally, we obtain samples of the approximate posterior distribution using an algorithm of the family of \textit{Markov chain Monte Carlo} (MCMC) methods.
%
% \end{frame}

@@

\vspace{1em}
\begin{alertblock}{Hyperparameter selection}
  \vspace{0.1em}
  For \(b_0\) we use a computational approximation of the MLE; the rest of hyperparameters are set via \textit{cross-validation}.
\end{alertblock}
\metroset{block=fill}
\end{frame}

@@

  \begin{block}{Logistic log-posterior}
      \vspace{-1em}
    \begin{align*}
  \log \pi(\theta \mid \mathcal D_n) \propto {} & \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi_{X_i}(\alpha)\right)Y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{X_i}(\alpha)\right\}\right)\right]\\
  \quad &+ \frac{1}{2}\log |G_\tau| - p\log \sigma -\frac{1}{2g\sigma^2} (b - b_0)'G_\tau(b - b_0).
\end{align*}
  \end{block}
  They are approximated through \textit{Markov chain Monte Carlo} (MCMC).
\end{frame}

@@

  \begin{exampleblock}{The road ahead}
    \begin{itemize}
      \item Reduce the variance in predictions.
      \item Derive theoretical properties of the predictors.
      \item Package the Python algorithm and make it compatible with standard libraries (e.g. \textit{sklearn}, \textit{scikit-fda}).
    \end{itemize}
  \end{exampleblock}
